Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-17 p2/axioms.tex
New ordering method; hyperlinks
Towards a new proof

We write $\Img{S}$ to denote the image of the domain of the sequence $X$.
\end{mydef}

@@

@@

\begin{myax}\label{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
\end{myax}

@@

\begin{myax}\label{ax_directparent_breaks}
All other commands on a parent break every filesystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
\end{myax}

@@

>>>>

We write $\Img{S}$ to denote the image of the domain of the sequence $S$.
\end{mydef}

@@

% Rules
% -----

@@

\begin{myax}\label{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.

We can derive this from our filesystem model in the following way:
we know if $y\neq\vald$ then the commands will break every filesystem $\FS$
as either before or after $\czwnv$, $\FS(n)$ is not empty, and therefore
at its parent $\nn$ there must be a directory.
We also know if $[z]\neq[\empt]$ then the commands will again break every filesystem
as either before or after $\cxynnv$, $\FS(\nn)$ is not a directory.
\end{myax}

@@

\begin{myax}\label{ax_directparent_breaks}
All other commands on a parent break every filesystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.

The derivation of this \namecref{ax_directparent_breaks} is similar to that of \cref{ax_directchild_breaks}.
We know if $y\neq\empt$ then the commands will break every filesystem $\FS$
as either before or after $\czwnnv$, $\FS(\nn)$ is not a directory, and therefore
every child of $\nn$ must be empty.
We also know if $[z]\neq[\vald]$ then the commands will again break every filesystem
as either before or after $\cxynv$, $\FS(n)$ is not empty.
\end{myax}

@@

% Independent commands
% --------------------

### 2016-10-17 p2/ordering.tex
Towards a new proof

In this section we describe a method with which all possible orderings of a
simple set of commands can be generated, but before doing so, let us consider the following lemma.

@@

\bigskip

\noindent
We can now define the method to generate possible orders of sets of commands.
What we expect is that if the commands can be applied to a filesystem
in some order without breaking it, then the method will yield all possible orders
in which they can be applied.

\begin{mydef}[$\ordersetsign$]
For a simple set or sequence $U\nequiv\cbrk$,
we use $\orderset{U}$ to denote the set of sequences
that are all permutations of the commands in $U$
for which the following holds:
for any two commands $\cxynv$ and $\czwmv$ in a sequence in $\orderset{U}$
where $\cxynv$ comes earlier in the sequence than $\czwmv$,
\begin{itemize}
\item if $n=\parent(m)$, then $y=\vald$
\item if $\parent(n)=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

In \cref{simple_reorder_equiv} we offer a formal proof that $\orderset{U}$ meets our expectations.
It is
an important property of simple sequences: namely, that
all of their valid reorderings are equivalent,
and all equivalent sequences are mere reorderings.
To prove this, first we prove the following:

@@

\begin{mylem}\label{equiv_simple_same_commands}
If two simple sequences $A$ and $B$ are equivalent
and do not break all filesystems ($A\equiv B\nequiv\cbrk$),
then they must contain the same commands.
\end{mylem}
\begin{proof}
We use a proof by contradiction. Let $A$ and $B$ be simple sequences
such that $A\equiv B\nequiv\cbrk$,
and $\FS$ be a filesystem that they do not break.

\begin{mylem}\label{simple_reorder_equiv}
For a simple sequence $S\nequiv\cbrk$,
\[ \orderset{S} = [S]_\equiv \cap \setssmp, \]
where $[S]_\equiv \cap \setssmp$ is the set of simple sequences equivalent to $S$.
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.

TODO:
Second, based on the algorithm we also know that $\orderset{S}$ contains
all possible orderings of $S$ as the only ordering between commands
it prescribes are when the commands in the reverse order would break every filesystem.

If, for example, in a component or sub-component
made up of construction pairs,
the command on the topmost node, $\cxynv$, was not before all other commands
in the component as prescribed by the algorithm,
then any resulting sequence would necessarily contain
a $\czwmv$ where $n\descendant m$, potentially followed by commands
on incomparable nodes, and then followed by $\cxynv$.
From \cref{ax_separate_commute} we know we can convert this sequence
to an equivalent sequence by swapping the commands on incomparable nodes with
$\czwmv$, which sequence would therefore have the commands on
$m$ and $n$ next to each other. Then from \cref{ax_distantrel_breaks,ax_directparent_breaks}
we know the sequence would break all filesystems
as a reversed construction pair is never a valid destruction pair.

We can prove that an order not generated by the algorithm
in the case of a component made of destruction pairs breaks all filesystems
in a similar way.

Also, from \cref{ax_separate_commute} we know
that the order between any other command pairs is not fixed in the sense
that if one order works, so will the other.

From these we know that all sequences in $\orderset{S}$
are equivalent, and that any sequence of the same commands outside it
necessarily breaks all filesystems.
We also know $S\in\orderset{S}$ as $S\nequiv\cbrk$
and therefore $T\in\orderset{S} \Rightarrow T\equiv S$.

From \cref{equiv_simple_same_commands} we know that 
if $T\equiv S$, then $T$ contains the same commands as $S$,
from which we see
that $T\equiv S\Rightarrow T\in\orderset{S}$.
\end{proof}

%% TODO https://en.wikipedia.org/wiki/Equivalence_class

It follows from the algorithm that:
\begin{mycor}\label{lemma:neighbor}
If $U$ contains commands on both $n$ and $\parent(n)$, then
there is a simple $U'\equiv U$ in which they are next to each other.
\end{mycor}

It follows from this that given a simple set,
we can simply assume that it has an order.
Accordingly, we will at times treat not only sequences as sets,
but simple sets as sequences.

>>>>

Let us therefore define $\ordersetsign$ which gives us all
possible orders in which a set of commands can be applied
to filesystems without breaking it
provided that the commands can be applied to a fileystem in some order.

\begin{mydef}[$\ordersetsign$]\label{def_orderset}
For a simple set or sequence $U$,
we use $\orderset{U}$ to denote the set of sequences
that are all permutations of the commands in $U$
for which the following holds:
for any two commands $\cxynv$ and $\czwmv$ in a sequence in $\orderset{U}$
where $\cxynv$ comes earlier in the sequence than $\czwmv$,
\begin{itemize}
\item if $n=\parent(m)$, then $y=\vald$
\item if $\parent(n)=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

In \cref{simple_reorder_equiv} we offer a formal proof that $\orderset{U}$ meets our expectations.
It is an important property of simple sequences: namely, that
all of their valid reorderings are equivalent,
and all equivalent sequences are mere reorderings.
To prove this, we will need the following results:

\begin{mylem}\label{order_is_only_possible}
For any simple sequence $S$ and its permutation $T$, 
\[ T\not\in\orderset{S} \Longrightarrow T\equiv\cbrk. \]
\end{mylem}
\begin{proof}
$T$ can only be excluded from $\orderset{S}$ if it contains commands on a node and its parent
which violate the conditions in \cref{def_orderset},
which, as simple sequences do not contain assertion commands,
entails that those commands placed next to each other in the same order
would not form a construction or destruction pair.

Informally speaking,
our proposition for simple sequences $T\nequiv\cbrk$ is equivalent to the statement that
two commands on a node and its parent must form
a construction or destruction pair even if they are not next to each other.
This is true because in simple sequences
there cannot be another command between either of the nodes involved,
so if the commands are incompatible, no command between them can change that.

% BACK_TO_THE_FS TODO
% TODO Review any proof depending on "nothing changes FS at n until...".
To prove our proposition formally, we reach back to our filesystem model.
Select two commands in $T$ that do not conform to \cref{def_orderset}, $\cxynv$ and $\czwmv$,
and split $T$ around them into three parts:
\[ T = T_0 \cc \cxynv \cc T_1 \cc \czwmv \cc T_2 \]
As $T$ is simple, there cannot be any commands on $n$ or $m$ in $T_1$,
and therefore $(T_0\cc\cxynv\cc T_1)\FS(n) = y$.
If $n=\parent(m)$, then we know $y\neq\vald$, and
(similarly to the reasoning in \cref{ax_directchild_breaks})
as either before or after $\czwmv$ which is not an assertion command,
$\FS(m)\neq\empt$, the latter command will break the filesystem.
If $\parent(n)=m$, then we know $y\neq\empt$, and
(similarly to the reasoning in \cref{ax_directparent_breaks})
as wither before or after $\czwmv$, $\FS(m)\neq\vald$,
the latter command will break the filesystem.
\end{proof}

@@

\begin{mylem}\label{equiv_simple_same_commands}
If two simple sequences $A$ and $B$ are equivalent
and do not break all filesystems ($A\equiv B\nequiv\cbrk$),
then they must contain the same commands.
\end{mylem}
\begin{proof}
% BACK_TO_THE_FS
We use a proof by contradiction. Let $A$ and $B$ be simple sequences
such that $A\equiv B\nequiv\cbrk$,
and $\FS$ be a filesystem that they do not break.

\begin{mylem}\label{simple_reorder_equiv}
For a simple sequence $S\nequiv\cbrk$,
\[ \orderset{S} = [S]_\equiv \cap \setssmp, \]
where $[S]_\equiv \cap \setssmp$ is the set of simple sequences equivalent to $S$.
\end{mylem}
\begin{proof}
First, we prove, by contradiction, that if a simple sequence $T\equiv S$, then $T\in\orderset{S}$.
We assume $T\equiv S$ and $T\not\in\orderset{S}$.
Then, from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv S\nequiv\cbrk$.

Next, we prove that if $T\in\orderset{S}$, then $T\equiv S$.
We proceed by induction.
Our base cases are $\emptyseq$, where $\orderset{\emptyseq}=\{\emptyseq\}$,
and one-long sequences, where this is trivially true.
In our induction step we assume that $T^*\in\orderset{S^*}\Rightarrow T^*\equiv S^*$ holds
for all sequences of length $i\geq 1$ or less.

Let us consider $T\in\orderset{S}$ where $T$ and $S$ are of length $i+1$.
We note that for any subsequence of $S$, $S_0$, which contains
only part of the commands in $S$ but in the same order as in $S$,
and for the corresponding subsequence of $T$, $T_0$, which
contains the same commands but in the order they are found in $T$,
$T_0\in\orderset{S_0}$ must hold as
if there were no command pairs in $T$ not conforming to \cref{def_orderset},
then there cannot be any in $T_0$.

We also know that due to \cref{ax_separate_commute,incomparable_is_independent}
$T$ can be rearranged into $T'\equiv T$ where
$T'=T_1\cc T_2\cc \cdots \cc T_k$ and the segments $T_j$ are pairwise independent,
and where we have the maximum number of segments possible.
$S$ can be converted similarly into $S'$, and
as because of \cref{equiv_simple_same_commands} $S$ and $T$ contain the same commands,
the segments of $S'$ will contain the same commands as those of $T$.
Let us therefore rearrange the segments of $S'$ so that the first segment
$S_1$ would correspond to (i.e. contain the same commands as) $T_1$, the second segment $S_2$ to $T_2$,
and so on.

The segments contain commands in the same order as the full sequences.
Therefore, from the observation abvout subsequences
we know for all $1\leq j\leq k$, $T_j\in\orderset{S_j}$,
and if $T'$ has at least two segments and so 
they are shorter than $i+1$,
from the induction hypothesis
we also know $T_j\equiv S_j$, from which we get $T\equiv S$.

We can assume $T'$ has at least two commands,
and in the remaining case, when $T'$ has only one segment,
we know all nodes that the commands in $T'$ apply to
are related via $\descendant$.
$T'$ contains the same commands as $S$, and
from \cref{connected_changes} we see that the nodes the commands apply to therefore form a tree in $\setn$.
We can therefore select the command on the topmost node (let it be $\cxynv$),
and we know $T'$ contains commands on some of its child nodes.
Let one of these be $\czwmv$.
Now mark the command $\czwmv$ and all commands in $T'$ that are applied to nodes that are descendants of $m$.
From \cref{ax_separate_commute} we know that all remaining commands in $T'$ (except $\cxynv$)
commute with the marked commands; therefore we can swap them to get $T''\equiv T'$
where marked commands are next to $\cxynv$:
\[ T'' = T_\mathrm{unmarked\ left} \cc T_\mathrm{marked\ left} \cc \cxynv \cc T_\mathrm{marked\ right} \cc T_\mathrm{unmarked\ right}. \]

TODO

First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.
\end{proof}

@@

TODO:
It follows from the algorithm that:
\begin{mycor}\label{lemma:neighbor}
If $U$ contains commands on both $n$ and $\parent(n)$, then
there is a simple $U'\equiv U$ in which they are next to each other.
\end{mycor}

TODO:
It follows from this that given a simple set,
we can simply assume that it has an order.
Accordingly, we will at times treat not only sequences as sets,
but simple sets as sequences.
